---
title: "Forecast daily bike rental demand using time series models"
date: "`r Sys.Date()`"
output: html_document
author: "Beatriz Moura"
---

# About Data Analysis Report

This page contains the report of the data analysis done for the project on forecasting daily bike rental demand using time series models in R. It contains analysis such as data exploration, summary statistics and building the time series models. The final report was completed on `r date()`.

**Data Description:**

This dataset contains the daily count of rental bike transactions between years 2011 and 2012 in Washington, D.C. bikeshare system with the corresponding weather and seasonal information.

**Data Source:** https://archive.ics.uci.edu/ml/datasets/bike+sharing+dataset

**Relevant Paper:**

Fanaee-T, Hadi, and Gama, Joao, 'Event labeling combining ensemble detectors and background knowledge', Progress in Artificial Intelligence (2013): pp. 1-15, Springer Berlin Heidelberg

# Task One: Load and explore the data

## Load data and install packages

This project depends on the `tidyverse` and the `timetk` packages. However, it was built on the `targets` framework.

Therefore, only the `targets` and the local packages will be loaded in this session.

```{r}
## Import required packages
library(targets)
tar_source()
```

The dataset was downloaded into the `data-raw` folder in a `.zip` extension.

```{r}
#| label: load data

```

## Describe and explore the data

```{r}

```



# Task Two: Create interactive time series plots

```{r}
## Read about the timetk package
# ?timetk


```




# Task Three: Smooth time series data

```{r}

```



# Task Four: Decompse and access the stationarity of time series data

```{r}

```



# Task Five: Fit and forecast time series data using ARIMA models

```{r}

```



# Task Six: Findings and Conclusions































